### 2019-07-09 justification_of_resources.tex
Justifcation of resources for edit

Justification of Resources

Staff, Directly Incurred Posts

PDRA Mullett. We request 100\% support for PDRA Mullett. A postdoctoral appointment is
essential here, as the level of technical skill required to run and analyse eye movement data is
high. As outlined in Staff Duties, Mullett will design the studies with the PI and lead on the
writing of the majority. Mullett will program the eye movement system and implement the
statistical analysis designed with the PI using the R and Matlab programming languages.
Mullett will also take a secondary role in the impact programme, and lead on the supermarket
collaboration.

PDRA at Caltech. We seek 100\% support for a 1-year PDRA at Caltech. The experimental
games eye tracking from Series 1, 3, 5, and 6 will be conducted at Caltech, together with a
subset of cross-lab replications.

RA. We request 25\% support for an RA. BSc or MSc level training in Psychology or
Behavioural Science will be sufficient. As outlined in Staff Duties, testing of participants one-
by-one, and testing 100 people is about 3 weeks of full time work. RA support will free
PDRA time for the more skilled writing and analysis.

Staff, Directly Allocated Posts

PI Stewart. We request 30\% support for Stewart. 15\%, half of this support, will be spent on
the laboratory programme, designing experiments and analysing data with the PDRA.
Stewart has developed a new statistical analysis for eye movement data (see two papers
forthcoming in Journal of Behavioural Decision Making) and this work will continue here.
The other 15\% is required for the ambitious impact work, which requires professorial level
and extensive liaison with the three current industry partners to negotiate access to data and
field trials, to set up access to three new partners, and to disseminate results back into
industry.

CI Chris Starmer. We will seek 10\% support for significant input from Experimental
Economist Prof. Starmer. Because of the interdisciplinary nature of the research, this
contribution is important for maximising the academic impact of the work beyond
Psychology.

Other Directly Incurred Costs

Computer resources. The computations required to estimate drift diffusion models and assess
model flexibility (e.g., with parameter space partitioning, Mullett & Stewart, 2015; Pitt,
Woojae, Navarro, & Myung, 2006) require serial search of model parameter spaces.
Compared to typical regressions which take fractions of a second, these methods take of the
order of days or weeks on, for example, a typical Intel 8-core i7-4770 CPU. We request
£2,000 for this dedicated computer machine.

Travel and Subsistence

We seek the costs of day visits to impact partners, all London based. We anticipate 6 visits per year x 2 people x 3 years. We will attend the SJDM conferences in 2016 and 2017. 2
people x 2 years (Vancouver and New Orleans) and the SPUDM conference (European city)
in 2017. These conferences are the most significant in the domain, giving coverage of
America and Europe. We seek the cost of a postdoc exchange with Prof. Colin Camerer at
Caltech. Caltech is the home of Camerer and Rangel labs, leaders in eye tracking economics.