### 2019-07-23 pathways_to_impact.tex
pathways to impact

\documentclass[11pt, a4paper]{article}

% Helvetica
\usepackage[scaled]{helvet}
\renewcommand\familydefault{\sfdefault} 
\usepackage[T1]{fontenc}

% 2cm margins
\usepackage[margin=2cm]{geometry}

\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[center]{caption}
\usepackage{hyperref}
\usepackage{placeins}
\usepackage{booktabs}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{threeparttable}
%\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}

\usepackage{apacite} 
%\usepackage{setspace}
%\doublespacing

\definecolor{green}{rgb}{0,1,0}
\newcommand{\NS}[1] {{\textcolor{green}{#1}}}
\newcommand{\TM}[1] {{\textcolor{orange}{#1}}}
\newcommand{\AT}[1] {{\textcolor{blue}{#1}}}

\begin{document}
\title{Understanding domestic abuse - Pathways to Impact}
\date{}
\maketitle

%See the section https://je-s.rcuk.ac.uk/Handbook/Index.htm#pages/GuidanceonCompletingaStandardG/CaseforSupportandAttachments/ESRCSpecificRequirements.htm on Pathways to Impact for ESRC.

Our pathway to impact is built upon the strong existing relationships we have with Wast Midlands Police (WMP) , and the college of policing. These existing relationships are complemented by our membership of the University of Warwick's Centre for Operational Policing Research (COPR), which provides access to an even wider established network within UK policing.

The research proposed here has been developed with direct input from officers and experts within WMP. This has allowed us to design the research questions that are relevant to current issues within WMP, and that can feed in to existing structures to accelerate the dissemination of findings, and changes in policy. To facilitate this, we have a day-to-day contact at WMP: Inspector Jenny Richards (Evidence Based Practice in Risk and Organisational Learning), and within the West Midlands Police and Crime Commissioner's Office, Alison Spence. As a result of this close collaboration, we have received support from senior levels within WMP including from Deputy Chief Constable Louisa Rolfe who is the National Police Chiefs Council lead for domestic abuse.

These relationships will enable us to work with WMP to import the insight from the research into their senior management team and also their day-to-day processes. This will be possible for each of the objectives and topics within our proposal:

1. The programme on ``predictors of serious harm'' will inform refinement of the THRIVE+ (threat, harm, risk, investigation, vulnerability, engagement, prevention and intervention) risk assessment model used for assessing calls and reports of domestic abuse. It will also provide a more general evidence base to help police develop policy.

2. The programme on the  ``decision to report'' links with police's design of the initial contact with victims (e.g., https://west-midlands.police.uk/your-options/domestic-abuse). It will also help the police to develop more efficient and targeted intervention and awareness campaigns.

3. The programme on ``police mis-recording of abuse'', is of direct interest to their data quality and reporting, as well as identifying any potential shortcomings in the way they deal with reports from particular types of victims.

4. The programme on ``environmental factors influencing abuse'' can be used to help predict which local populations are most at risk of abuse in the West Midlands. It can also give clues as to likely triggering events, such as economic shocks, or changes to benefits, and large events such as sports championships.

In addition to working with WMP, our links to the National Police Chiefs council, and collaboration with the College of Policing, provide pathways for impact across all forces. At the College of Policing (situated just a few miles from Warwick campus) we will work within the framework of the \href{https://www.app.college.police.uk/app-content/major-investigation-and-public-protection/domestic-abuse/introduction/) and the revision of the existing DASH risk assessment tool for domestic abuse. (https://whatworks.college.police.uk/Research/Documents/DA_risk_assessment_pilot.pdf}{Authorised Professional Practice} on domestic abuse, released in 2015. We will work initially with David Tucker, Head Crime and Criminal Justice at the College of Policing.

We will meet regularly with our contacts at WMP and the college of policing throughout the grant. This will enable us to adapt to current questions that arise during the period of the proposal. It also allows us to react quickly to any unanticipated issues that arise as a result of our findings, such as following up on unexpected results to identify underlying causes. It also ensures that there is an ongoing dialogue about the data being used, and provides an opportunity to alert the police to any data quality issues.

In the second year of the proposal we will hold a 1-day impact summit at Warwick Business School's London Shard venue. This will be funded by the Warwick Business School Impact Fund, and potentially the ESRC Impact Acceleration Account. We will host the National Police Chiefs' Council lead for Domestic Abuse, Deputy Chief Constable Louisa Rolfe (who is also at West Midlands Police), and Members of Parliament from the Home Affairs Committee domestic abuse inquiry. The summit will be a 50/50 academic / practitioner mix. We will contact, early in the project, [Insert list of charities in this sector, Women's Aid, Refuge, Safe Lives, National Centre for Domestic Abuse] and visit them at their offices to present the programme of research and involve them as we conduct the initial analysis.

In addition to the academic journal outputs, we will present the findings in a series of thought leadership pieces for dissemination across the police and third-sector contacts. We will submit these reports to the College of Police What Works database and log them on the \href{https://whatworks.college.police.uk/Research/Research-Map/Pages/Research-Map.aspx}{Research Map}.

Although a high-profile ``launch'' event at the Shard will attract attention, it is the continuous week-on-week contact with our impact partners that we think will deliver the legacy impact for this project.

Within Lloyds Bank, we will work with Martin King, Head of Customer Vulnerability. Warwick already has a relationship with King's team in CI Stewart's work on the effects of gambling, which was completed using micro-level transaction data from millions of account holders. Key issues here are work on coercion and control, and having sufficient financial resilience to escape an abusive relationship and being able to make the decision to report an abuser. Our day-to-day contact at Lloyds is Dr David Leake, Head of the Behavioural Insights Team. We will present the findings from all four programmes to the Lloyds Vulnerability Team in six-monthly briefings, and work to integrate insight from the research into the Lloyds environment.

%Events:

%Impact summit (1 day at the Shard) \TM{We haven't budgetted anything for this...}
%Visits to key charities
%Visits to West Midlands Police
%Presentation to NPCC
%Data sprint session with WMP Behavioural Insights Team?

PI Mullett will spend half of his time of the project leading the impact activity and takes overall responsibility of delivering the impact programme. Mullett has worked extensively on the impact programme from the ESRC funded ``Accumulating to Choose'' programme, with the Financial Conduct Authority. CI Stewart will also spend half of his time of the project on the impact activity. Stewart has significant experience of impact acitivites in the domain of financial decision making with Lloyds Banking Group, Barclays Bank, and the UK Finance Association. Postdoc Trendl will be fully involved in the impact programme. This is an important part of the professional development for an early career researcher, and will allow her to establish a network of independent contacts.

We will apply to the Warwick ESRC Impact Acceleration Account and to Warwick Business School Impact Fund to continue the impact programme beyond the end of the funded period.

%
%
%
%
%\section{Old Accumulating to Choose Pathways to Impact}
%
%Pathways to Impact
%We will target those in industry and public policy with interest in economic decision making.
%While the focus of this programme is on informing micro-level theory of economic choice,
%we will develop the implications for regulation and for the competitiveness of UK financial
%industry. Our strategy will be to translate the laboratory work detailed in the Case for Support
%into the field with a series of randomised control trials, detailed below. To deliver significant
%impact it is essential for us to bridge the divide from the laboratory by working on live
%problems for the regulator and industry using real consumers making real decisions.
%Since their inception in 2012 the Financial Conduct Authority (the UK regulator for all
%financial products, like savings accounts, pensions, mortgages, credit cards, and insurance)
%have placed significant emphasis on behavioural regulation. This includes “...getting a better
%understanding of why consumers act in the way they do, so we can adapt our regulation to
%their [consumers’] common behavioural traits” (Martin Wheatley, outgoing Chief Executive,
%FCA). Indeed, the first publication from the FCA made the case for using behavioural
%economics in regulation (Erta, Hunt, Iscenko, Brambley, 2013). Recently Stewart has worked
%closely with the new Behavioural Economics and Data Science Team at the FCA, led by Dr
%Stefan Hunt, on the psychology of structured deposits (Hunt, Stewart, \& Zaliauskas, 2015).
%The FCA are at the beginning of a project on the complexity of financial products. The
%pathway to impact within the FCA will be via regular meetings between members of the FCA
%Behavioural and Data Science Team, and Stewart and Mullett (most recently in August 2015).
%Ultimately, as the data will drive public policy change regarding presentation and advertising
%of financial products, with the creation of regulations grounded in behavioural science. The
%outcome of this engagement will be the design of randomised control trials in collaboration
%with the FCA and a customer-facing financial services provider using real consumers making
%real decisions. Dissemination via the FCA working paper series will have more significant
%reach than academic publication alone. Please see the letter of support from the Financial
%Conduct Authority.
%Warwick has a formal collaboration with the consumer group Which?, and Stewart has
%worked closely with Alex Chesterfield, Head of Behavioural Insights, and team members
%Sam Himmelweit, Harriet Patterson, and Phillida Cheetham. The formal collaboration covers
%data sharing, exploitation, and publication rights, etc., and we have several existing and
%completed projects (e.g., on credit card repayment behaviour and expectant mums’ birth
%choices). Early insights from this work will feed into Which?’s project on the complexity of
%financial products. They key route will be the monthly joint meetings between Warwick and
%Which? Please see the letter of support from Which? Again, we would seek to implement
%insight from these grants in a practical financial decision domain as part of Which?’s
%consumer testing programme, and thus we will identify a practical application of the drift
%diffusion model to a consumer choice of value to which and implement a field trial. Likely
%candidates are: (a) the development of product information sheets for mortgages, savings
%accounts, pensions, etc., within the context of providing behavioural nudges, and (b)
%information provision for telecoms choices.
%The UK Cards Association is the industry body for all credit card providers in the UK.
%Stewart (and Gathergood, Economist, Nottingham) have a significant existing collaboration,
%with the sharing of millions of credit card statements between about 10 of the largest card
%companies and Warwick and Nottingham Universities. Contacts are Paul McCarron, Head of
%Cards and Fraud Control, and Richard Koch, Head of Policy. Stewart contributed evidence to
%the 2009 White Paper reviewing of the credit industry. Of key interest is (a) consumers’
%
%decisions around credit card repayments, and their integration of information from monthly
%statements and (b) consumers choice of card during switching and how cards are evaluated.
%The former fits well with the evaluation work; the latter fits well with the choice work.
%Dissemination of research findings will be via presentation to the Current Affairs working
%group, with representatives from every UK card provider.
%We will also work closely with the Economic and Social Research Institute, an independent
%research institute funded mainly by the Irish Government. Our key contact is Dr Peter Lunn,
%who has expertise in the application of decision making and consumer psychology research to
%policy and regulation.
%These routes to impact are time consuming. They would be led by the PI, with significant
%involvement from the CI to distribute the load. We anticipate that 15% (i.e., half) of PI time
%will be spent pursuing impact. These contacts will also add value for the PDRA and CI as
%they can be taken to build research collaborations beyond the grant, independently of the PI.
%We have requested funding to develop the field trials over the course of the grant, and in
%previous experience the ability to make modest commitment to a larger programme of
%randomised control trials enhances engagement.
%Beyond these three existing routes to impact, during the course of the grant we will seek to
%extend our contacts and influence to (a) one major UK supermarket, (b) one major UK
%insurance company, and (c) one major bank.
%Table 1. Translational field studies.
%Translational Field Study
%
%Anticipated Design
%
%1. Financial Conduct
%Authority
%
%A representative sample of 2,000 investors in structured
%deposits, payday loans, or other complicated financial
%products.
%
%2. Which?
%
%Presenting multi-attribute choice options to 2,000
%consumers switching broadband and or expectant mums.
%
%3. UK Cards
%
%A representative sample of 2,000 cardholders making (a)
%direct debit setup decisions or (b) monthly repayment
%decisions.
%
%Finally, the opportunity to work with real consumers making real decisions with non-trivial
%economic consequences (maybe tens of pounds with broadband, hundreds of pounds with
%credit cards, and thousands of pounds with structured deposits of mortgages, and hours of
%pain with expectant mums) is of significant academic value. The issue of incentivized choice
%is critical in Experimental Economics (though less so in Psychology), and this impact work
%extends the lab work into high stakes choice.
%
%\section{Old Perfect Cop Pathways to Impact}
%
%
%1
%
%Pathways to Impact
%This research is likely to deliver impact in recruitment of new officers, identifying outcomes
%associated with the best management, reductions in the rates of sickness, reductions in
%complaints and better responses to complaints, and matching officers to roles. In particular,
%matching responses from our Online Lab to the police data adds value to those data. For
%example, if, as we suspect, conscientiousness does indeed correlate with rates of sickness
%and complaints, this would allow forces to understand sickness and complaints as, at least in
%part, a function of individual differences in personality interacting with the local context in
%which officers and staff are placed. And if well-being is associated with the quality of
%management, this is a first step in developing mechanisms for improving the well-being of
%officers and staff. Below we detail the pathways to these potential impacts.
%The police support this research and have provided their data and access to officers and
%staff without imposing constraints on the academic research programme. We have met with
%and discussed the project with all of the officers and staff named here (and see meeting
%dates on p. 2 of the Case for Support). At the MPS, this research project reports to the Total
%Professionalism Board, which is chaired by our force lead Assistant Commissioner Helen
%King. The Board is responsible workforce development and recruitment strategy. We have
%already made presentations to the Board (April 2015, March 2016) and received their
%backing. By integrating this project with the Board’s remit, we have official and long-term
%support for the project. At Nottinghamshire the project reports to Chief Constable Sue Fish.
%At States of Jersey the project reports to Chief Officer Mike Bowron.
%The findings of this study have relevance to both metropolitan and county forces throughout
%England, Wales and Northern Ireland (whom all operate under a common Doctrine and
%Professional Standard). At the local level we will give at least six monthly feedback. For each
%force, each Chief Constable and Police and Crime Commissioner pairing has independent
%authority over their area (or MPS Commissioner / London Mayor pairing), and we will use
%these meeting to develop interventions and analysis around the five themes below. At the
%national level, the MPS is the largest police force in the UK and is a leader in practice. The
%route to national-level impact is the National Police Chiefs’ Council (business area lead is
%Chief Constable Giles York; Chief Constable Sue Fish is our NPCC sponsor) and through
%the College of Policing who inform the National Approved Professional Practice (Rachel
%Tuffin, Director, Knowledge, Research, and Education). At the international level, the high
%regard of British policing provides a reference point for international standards. We have
%already made links with the Chicago Citizens’ Police Data Project, Harvard Sociology, and
%Chicago Police Department, and with New York Police Department via Warwick’s Institute for
%Sustainable Cities. National and international level feedback will be via Warwick’s Centre for
%Operational Policing Research, co-directed by PI Stewart with CI Hodgson.
%Feedback to forces will be structured against the following themes, which have been
%highlighted in discussions with senior officers, who will lead the implementation within their
%forces. Feedback points are every six months, as agreed with the officers and staff named
%below.
%Public Desire. The research will deliver a list of psychological indicators of good
%performance, reduced complaints and misconduct, and reduced physical and mental ill
%health. These measures will inform the recruitment process and the selection of officers and
%staff for specialist roles in the MPS. But changing recruitment processes represent a shift
%from the implicit objective of police in the image of the public they serve. For example, “We
%can’t have a police force representative of a London which has an ethnic minority population
%of 55% when only 10% of its officers are from an ethnic minority” (Lammey, 13 January
%2014; the 55% is probably a little too high here). Many forces use positive recruitment of
%BME officers, and the MPS used residence requirements in 2013/14. How have recent
%changes in recruitment changes the psychological attributes selected for? There is certainly
%a political imperative for this directive, but little empirical evidence that the public wish this as
%
%2
%a determining factor for officer or staff recruitment. To this end, we will complement the
%behavioural science evidence base with a survey of the public to measure the properties of
%officers they desire (with Chief Constable Sue Fish, Nottinghamshire; Assistant
%Commissioner Martin Hewitt, MPS; and the (London) Mayor’s Office for Police And Crime).
%We will use a market research company to measure, from a representative sample of the UK
%population, their views on the age, sex, race, ethnicity, religious affiliation, psychological
%traits. Our objective is to start and develop a public dialogue.
%Management. What is the impact of performance development review gradings on both
%individual behaviours and those of their subordinates (Deputy Commissioner Craig Mackey
%and Chief Superintendent Robert Jones). That is, what outcomes in are associated with
%better line managers? We have already found that the highest scoring line managers have
%subordinates with a 20% reduction in the incidence of sickness. We have also identified
%significant variability across the Met within the sickness and complaints measures, with
%some boroughs having half the rates of sickness and complaints of others, and some
%boroughs having large changes within the 5-year period of the data. The mechanism for
%impact within the MPS is the HR change programmes (Robin Wilkinson, HR Director).
%Absenteeism and complaints. Using the data we already have, we see a very strong
%association between sickness and complaints. Among officers who have no complaint in the
%5-year period, 63% of officers have one or more episode of sickness. But among officers
%who have a complaint, 84% have one of more sickness. Further, econometric panel data
%analysis shows that there is almost no association between previous complaints and current
%sickness or vice versa: Complaints do not appear to cause sickness, and sickness does not
%put officers at greater risk of a complaint. The logical conclusion from this work is that there
%are strong individual difference predictors of these measures—and we have
%conscientiousness and agreeableness from the online lab as prime candidates. So key
%questions are why are complaints and sickness associated (Chief Constable Sue Fish,
%Nottinghamshire)? Which individual differences predict complaints and sickness (Chief
%Superintendent Robert Jones)?
%Right for the role? Do certain types of behavioural attributes (personality, risk attitudes)
%coalesce in specific police functions, such as specialist operations or investigations (Chief
%Officer Mike Bowron, States of Jersey Police)?
%Bad cops or bad circumstances? Are there career hotspots for misconduct or serious
%substantiated complaints (Deputy Commissioner Craig Mackey). Can we provide early
%intervention warnings for officers at greatest risk of misconduct? The mechanism for impact
%is through Professional Standards (Deputy Assistant Commissioner Fiona Taylor), to help
%design career paths that avoid potentially potent placements or role succession that
%materially increase the risk of misconduct.
%We will also offer a series of four 1-day Behaviour Masterclasses to Senior Officers'
%departments in the final year of the project. Classes will be advertised via the College of
%Policing, the National Police Chief’s Council, and Warwick’s Centre for Operational Policing
%Research. We will target the chief officers responsible for policy and implementation within
%their force, and members of their teams. Classes will be co-led by Stewart or Ritchie, and an
%officer. We will use feedback to shape the further development of the research.
\newpage

\bibliographystyle{apacite}
\bibliography{domesticabuse_refs}

\end{document}