### 2019-07-23 accumulating_to_choose_interim_report.tex
Accumulating to choose interim report

Progress report for ESRC grant ES/N018192/1 ``Accumulating to choose''

Accumulating to choose is an ESRC grant exploring how evidence is accumulated during decision making. It is a collaboration between psychologist and PI Prof. Neil Stewart, and economists Prof. Colin Camerer and Prof. Chris Starmer.

Named postdoc Dr Tim Mullett began working on the grant on 1 January 2017.Caltech postdoc Devdeepta Bose was hired July 2018, with matched funding from Caltech extending his contract for a second year.  Mullett was hired as an assistant professor in September 2018, but continues to work on the grant. Replacement Dr Charlotte Edmunds joined the project A no-cost extension was granted until 30th June 2020 (in March 2019). Thus the project is now in its third of three funded years.

During the grant over 18 published papers carry the grant number. Core outputs are:

Mullett, T. L., Brown, G. D. A., \& McDonald, R. (2019). Cooperation in public goods games predicts behaviour in incentive-matched binary dilemmas : evidence for stable pro-sociality. Economic Inquiry.

Mullett, T. L., Brown, G. D. A., Fincher, C. L., Kosinski, M. \& Stillwell, D. (2019). Individual-level analyses of the impact of parasite stress on personality: Reduced openness only for older individuals. Personality and Social Psychology Bulletin.

Bhatia, S., \& Mullett, T. L. (2018). Similarity and decision time in preferential choice. Quarterly Journal of Experimental Psychology.

Alempaki, D., Canic, E., Mullett, T. L., Skylark, W. J., Starmer, C., Stewart, N., \& and Tufano, F. (2018). Re-examining how utility and weighting functions get their shapes: A quasi-adversarial collaboration providing a new interpretation. Management Science.

Mullett, T. L., Smart, L., \& Stewart, N. (2018). Blackbird's alarm call or nightingale's lullaby? The effect of tweet risk warnings on attractiveness, search, and understanding. Financial Conduct Authority Occasional Papers in Financial Regulation, 47. https://www.fca.org.uk/publication/occasional-papers/occasional-paper-47.pdf PDF 
Abstract

Mullett, T. L., Smart, L., \& Stewart, N. (2018). The effect of risk warning content for contract for difference products. Available at SSRN: https://ssrn.com/abstract=3231616

Noguchi, T., \& Stewart, N. (2018). Multialternative decision by sampling: A model of decision making constrained by process data. Psychological Review, 125, 512-544. doi: 10.1037/rev0000102

We also have a series of working papers currently under review or in progress. Those most relevant include:

Edmunds, C., Mullett, T. L., \& Stewart, N. No accumulation in value based choice.

Stewart, N., Mullett, T.L., \& Canic, E. On the futility of estimating utility functions: Why the parameters we measure are wrong, and why they do not generalize.

Mullett, T.L., \& McDonald, R. Visual Attention in Health Judgements and Choices.

McDonald, R., Mullett, T.L., \& Tuschia, A. Perceptions of composite attributes in judgements of EQ5D.

Brown, G.D.A., Walasek, L., Mullett, T.L., Fincher, C., Stillwell, D., \& Kosinski, M. Pathogen Stress and Politics: Changing Ecological Origins of Ideology in the USA.

Mullett, T.L., \& Stewart, N. Don’t fixate on interactions, attend to main effects: Attentional models of choice are miss-specified.

Mullett, T.L., \& McDonald, S. How People Choose Universities: The Impact of TEF and Variable Fees.

Walasek, L., Mullett, T.L., \& Stewart, N. A Meta-Analysis of Loss Aversion in Risky Contexts. Available at SSRN: https://ssrn.com/abstract=3189088

Mullett, T.L., \& Stewart, N. No evidence for the payday effect in alcohol related crimes and hospital admissions.

Mullett, T.L. Stewart, N., \& Tunney, R.J. Time course of attention patterns during multi-attribute choice reveals attribute weighting and preference formation in a two stage process.

We have our final round of exchange visits soon, with Bose visiting in August and Mullett visiting Caltech in September.

Data collection is nearly complete. We have one final set of experiments to collect data for in autumn 2019.